---
title: "EDA-challenge.Rmd"
author: "Sydney Jansen"
date: "2024-02-06"
output: html_document
editor_options: 
  chunk_output_type: console
---
##Step 1
```{r }
# Read the CSV file and assign the resulting dataframe to a variable
library(tidyverse, dplyr)
f<-("C:/Users/Wildc/Downloads/data-wrangling.csv")
d<-read.csv(f, sep = ",", header = TRUE, stringsAsFactors = FALSE)
names(d)

# Now use head() on the dataframe to display the first few rows
df<-data
head(df)
#create new variable for body size dimorphism: ratio of average male to female body mass.
df <- df |>
  mutate(BSD = Body_mass_male_mean / Body_mass_female_mean)

# Create sex_ratio as the ratio of the number of adult females to adult males
# Here won't recognize variables
df <- df |>
  mutate(sex_ratio = AdultFemale/AdultMales)

# Create DI (Defensibility Index) as the ratio of day range length to the diameter of the home range

df <- df |>
  mutate(
    home_range_diameter = 2 * sqrt(HomeRange_km2 / pi),
    DI = DayLength_km / home_range_diameter
  )

library(ggplot2)             
str(df)

p <- ggplot(data = df, aes(x = Family, y = DI)) +
  geom_boxplot(na.rm = TRUE)
print(p)

  
geom_point() +
  labs(title = "Day Range Length vs. Time Spent Moving",
       x = "Time Spent Moving",
       y = "Day Range Length",
       color= Species) +
  theme_minimal()


ggplot(df, aes(x = DI, y = DayLength_km, color = Family)) +
  geom_point() +
  labs(title = "Day Range Length vs. Time Spent Moving",
       x = "Time Spent Moving",
       y = "Day Range Length",
       ) +
  theme_minimal()


 
```

## Do species that spend more time moving travel farther overall? 
  No
How about within any particular primate family?
  Families that travel for longer period travel farther (Cebidae) 
Should you transform either of these variables?
  No

```{r }
ggplot(df, aes(x = DI, y =  MeanGroupSize, color= Family)) +
  geom_point() +
  labs(title = "Day Range Length vs. Time Spent Moving",
       x = "Time Spent Moving",
       y = "Time Group Size") +
  theme_minimal()

```

## Do species that live in larger groups travel farther overall?
  Larger groups, less moving
How about within any particular primate family?
  Cebidae spends more time moving with a large group
Should you transform either of these variables?
  No
```{r}
#overall

ggplot(df, aes(x = BSD, y= Canine_Dimorphism)) +
  geom_point() +
  labs(title = "body size dimorphism and canine size dimorphism",
     x = "BSD",
       y = "Canine") +
  theme_minimal()
#family
ggplot(df, aes(x = BSD, y= Canine_Dimorphism , color= Family)) +
  geom_point() +
  labs(title = "body size dimorphism and canine size dimorphism",
     x = "BSD",
       y = "Canine") +
  theme_minimal()
```

#Do taxa with greater size dimorphism also show greater canine dimorphism?
Yes

```{r}
df <- df |>
  mutate(diet_strategy = case_when(
    Fruit    > 50 ~ "frugivore",
    Leaves > 50 ~ "folivore",
    TRUE ~ "omnivore"
  ))


# Plotting group size by dietary strategy
ggplot(df, aes(x = diet_strategy, y = MeanGroupSize, fill = diet_strategy)) +
  geom_boxplot() +
  labs(title = "Group Size by Dietary Strategy",
       x = "Dietary Strategy",
       y = "Group Size") +
  theme_minimal()

```
#Do frugivores live in larger groups than folivores?
No
```{r}
library(dplyr)
f<-("C:/Users/Wildc/Downloads/data-wrangling.csv")
d<-read.csv(f, sep = ",", header = TRUE, stringsAsFactors = FALSE)
names(d)

d <- d |>
  mutate(Binomial = paste(Genus, Species)) |>
  select(Binomial, Family, Brain_Size_Species_Mean, Body_mass_male_mean) |>
  group_by(Family) |>
 summarise( 
    Avg_Brain_size_species_mean = mean(Brain_Size_Species_Mean, na.rm = TRUE),
    Avg_Body_mass_male_mean = mean(Body_mass_male_mean, na.rm = TRUE)
)|>
  arrange(Avg_Brain_size_species_mean)


print(d)
```