---
title: "EDA-challenge.Rmd"
author: "Sydney Jansen"
date: "2024-02-06"
output: html_document
editor_options: 
  chunk_output_type: console
---
##Step 1
```{r }
# Read the CSV file and assign the resulting dataframe to a variable
library(tidyverse, dplyr)
f<-("C:/Users/Wildc/Downloads/data-wrangling.csv")
d<-read.csv(f, sep = ",", header = TRUE, stringsAsFactors = FALSE)
names(d)

# Now use head() on the dataframe to display the first few rows
df<-data
head(df)
#create new variable for body size dimorphism: ratio of average male to female body mass.
df <- df |>
  mutate(BSD = Body_mass_male_mean / Body_mass_female_mean)

# Create sex_ratio as the ratio of the number of adult females to adult males
# Here won't recognize variables
df <- df |>
  mutate(sex_ratio = AdultFemale/AdultMales)

# Create DI (Defensibility Index) as the ratio of day range length to the diameter of the home range
##
df <- df |>
  mutate(
    home_range_diameter = 2 * sqrt(HomeRange_km2 / pi),
    DI = DayLength_km / home_range_diameter
  )

library(ggplot2)             
str(df)

p <- ggplot(data = df, aes(x = Family, y = DI)) +
  geom_boxplot(na.rm = TRUE)
print(p)

  
geom_point() +
  labs(title = "Day Range Length vs. Time Spent Moving",
       x = "Time Spent Moving",
       y = "Day Range Length") +
  theme_minimal()

#I cannot figure out how to scale down the plot ggplot(df, aes(x = Move, y = DayLength_km, color = Species)) +
ggplot(df, aes(x = DI, y = DayLength_km)) +
  geom_point() +
  labs(title = "Day Range Length vs. Time Spent Moving",
       x = "Time Spent Moving",
       y = "Day Range Length") +
  theme_minimal()
 
```

## Do species that spend more time moving travel farther overall? 
  No
How about within any particular primate family?
  Families that travel for longer period travel farther (Cebidae) 
Should you transform either of these variables?
  I need to transform the species to fit the graph

```{r }
p <- ggplot(data = d, aes(x = log(Body_mass_female_mean), y = log(Brain_Size_Species_Mean),
    color = factor(Family)))  # first, we build a plot object and color points by Family
# then we modify the axis labels
p <- p + xlab("log(Female Body Mass)") + ylab("log(Female Brain Size)")
# then we make a scatterplot
p <- p + geom_point(na.rm = TRUE)
# then we modify the legend
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
# and, finally, we plot the object
p

```

## Do species that live in larger groups travel farther overall? 
How about within any particular primate family?
Should you transform either of these variables?

```{r pressure, echo=FALSE}
plot(pressure)
```

#Do taxa with greater size dimorphism also show greater canine dimorphism?
